Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-05-31 Documents/M1/realizableFermionicM1/sections/appendix.tex
Added 'Absolute Stability' subsectionto IMEX section.
Work on imex section.  Replaced all hu_et_al_2017 with hu_etal_2018

A second-order accurate, positivity-preserving IMEX scheme of type $A$ (the matrix $A$ is invertible) was given in \cite{hu_etal_2017}.  
We refer to this scheme as IMEX PA2.  
For this scheme, the non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 0.7369502715, \nonumber \\
  \tilde{a}_{31} &= 0.3215281691, \quad \tilde{a}_{32} = 0.6784718309, \nonumber \\
  a_{11} &= 0.6286351712, \nonumber \\
  a_{21} &= 0.2431004655, \quad a_{22} = 0.1959392570, \nonumber \\
  a_{31} &= 0.4803651051, \quad a_{32} = 0.0746432814, \quad a_{33} = 0.4449916135. \nonumber
\end{align}
The coefficient in the correction step is $\alpha = 0.2797373792$ and the CFL constant is $c_{sch} = 0.5247457524$.
This scheme is globally stiffly accurate (GSA), so that $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

@@

\newpage 
\section{Construction of PARSD}

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2017}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i},\\
 \text{(2nd-order accuracy)} \qquad\ &\sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0 ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PARSD scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.

>>>>

A second-order accurate, positivity-preserving IMEX scheme of type $A$ (the matrix $A$ is invertible) was given in \cite{hu_etal_2018}.  
We refer to this scheme as IMEX PA2.  
For this scheme, the non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 0.7369502715, \nonumber \\
  \tilde{a}_{31} &= 0.3215281691, \quad \tilde{a}_{32} = 0.6784718309, \nonumber \\
  a_{11} &= 0.6286351712, \nonumber \\
  a_{21} &= 0.2431004655, \quad a_{22} = 0.1959392570, \nonumber \\
  a_{31} &= 0.4803651051, \quad a_{32} = 0.0746432814, \quad a_{33} = 0.4449916135. \nonumber
\end{align}
The coefficient in the correction step is $\alpha = 0.2797373792$ and the CFL constant is $c_{sch} = 0.5247457524$.
This scheme is globally stiffly accurate (GSA), so that $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

@@

\newpage 
\section{Construction of PARSD}
\label{app:PARSD}

It can be shown that no 2-stage positivity-preserving IMEX exist \cite{hu_etal_2018}.
Hence we considered 3-stage IMEX scheme.
There are two kinds of implicit DIRK scheme, Type A and Type ARS.
Here we discuss the simpler case, Type ARS, first.
The Butcher tableau for ARS type IMEX scheme is
\begin{equation}
  \begin{array}{c | c c c}
  	         0           & 0                    & 0                    & 0                    \\
  	\tilde{c}_{2} & \tilde{a}_{21}       & 0                    & 0                    \\
  	\tilde{c}_{3} & \tilde{a}_{31}       & \tilde{a}_{32}       & 0                    \\ \hline
  	                     & \tilde{w}_{1} & \tilde{w}_{2} & \tilde{w}_{3}
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0            & 0            & 0            \\
  	c_{2} & 0            & a_{22}       & 0            \\
  	c_{3} & 0            & a_{32}       & a_{33}       \\ \hline
  	             & w_{1} & w_{2} & w_{3}
  \end{array}.
\end{equation}
The properties a desired IMEX scheme should have are: 2nd-order accurate in the streaming, globally stiffly accurate, positivity-preserving, diffusion accurate and an affordable time-step restriction.
Then the problem that find such a scheme can be rewritten as an optimization problem:\\
Find $(\tilde{a}_{21}, \tilde{a}_{31}, \tilde{a}_{32}, \tilde{w}_{1}, \tilde{w}_{2}, \tilde{w}_{3}, a_{22}, a_{32}, a_{33}, w_{1}, w_{2}, w_{3})$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ \dfrac{c_{20}}{\tilde{c}_{20}}, \dfrac{c_{30}}{\tilde{c}_{30}}, \dfrac{c_{32}}{\tilde{c}_{32}}\}\\
 \text{under these constraints:} \qquad & \\
 \text{(GSA)} \qquad\ &\tilde{w}_{i} = \tilde{a}_{3i}, \quad w_{i} = a_{3i}, ~\text{for} ~ i = 1,2,3\\
 \text{(2nd-order accuracy)} \qquad\ &\sum_{i=1}^{3}  \tilde{w}_{i} = 1, \quad \sum_{i=1}^{3}  \tilde{w}_{i}\tilde{c}_{i} = 1/2, \\
 \text{(positivity-preserving)} \qquad & a_{ii} > 0, c_{i0} \geq 0, \tilde{c}_{i0} \geq 0, ~\text{for} ~ i = 2,3 \\
 & c_{32} \geq 0, \tilde{c}_{32} \geq 0, \\
 \text{(diffusion accurate)} \qquad & e^{T}_iA^{-1}\tilde{A}e = 1, ~\text{for} ~ i = 2,3, \\
 \text{(affordable)} \qquad & c_{sch} > 0.
\end{align*}
Imposing those equality constraints (including all the constraints exclude those from the positivity-preserving and affordable requirements), the Butcher tableau can be simplified as:
\begin{equation}
  \begin{array}{c | c c c}
  	     0       & 0            & 0 & 0 \\
  	\frac{1}{2x} & \frac{1}{2x} & 0 & 0 \\
  	     1       & 1-x          & x & 0 \\ \hline
  	             & 1-x          & x & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	     0       & 0 & 0            & 0 \\
  	\frac{1}{2x} & 0 & \frac{1}{2x} & 0 \\
  	     1       & 0 & 1-y          & y \\ \hline
  	             & 0 & 1-y          & y
  \end{array}.
\end{equation}
And the rest inequalities and the problem were left to be: \\
Find $(x,y)$ such that
\begin{align*}
\text{maximize} ~ c_{sch} = &\min \{ 2x, \dfrac{1-2x(1-y)}{y-x}, 2(1-y)\}\\
 \text{under these constraints:} \qquad & x > 0, ~ 1 \geq y > 0, \\
 & x \leq y, ~\dfrac{1}{2x} \geq 1-y.
\end{align*}
A solution family is $x=\frac{1}{2}$, $y=\frac{1}{2} + \epsilon$ with $\epsilon \in [0, \frac{1}{2})$ and $c_{sch} = 1 - 2\epsilon$.
The PARSD scheme follows.
It's clear that $\epsilon = 0$ and $c_{sch} = 1$ is the most optimized solution.
